---
title: "iSC.MEB: installation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{iSC.MEB: installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install the iSC.MEB
This vignette provides an introduction to the R package `iSC.MEB`, where the function `iSC.MEB` implements the model `iSCMEB`, a integrate spatial clustering analysis method with hidden Markov random field using empirical Bayes.
The package can be installed with the following command from [Github](https://github.com/XiaoZhangryy/iSC.MEB):

`install.packages('devtools')`
`devtools::install_github("XiaoZhangryy/iSC.MEB")`

The package can be loaded with the command:
```{r  eval=TRUE}
library("iSC.MEB")
```


## Setup on Linux or MacOS system
For running big data, users can use the following system command to set the C_stack unlimited in case of `R Error: C stack usage is too close to the limit`.
```{Linux}
ulimit -s unlimited
```